---
title: "Module 3"
author: "Harrison Leduc"
date: "2024-01-19"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{module3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette covers Unit 1 Module 3. Tasks are numbered and in order. This vignette works with a newly implemented function, 'my_calc'.

Questions:
"Why shouldn’t you use the name “df” to name this object?"
  df could create a conflict with other package function names. For instance, 
 I believe there is a density function that is stated as df.
 Also df is a common 'junk name' for people creating data frames. 
 Something more specific should be used than DF, however, it is a step up 
 from df. df being a junk name means it can easily be overwritten and 
 lacks a description due to its vagueness. 
 
"Why is it okay to rename a column in DF “c”, but not to rename v3 as c?"
  Renaming a column in DF "c" only adjusts the DF. You should not rename v3 to c
 because v3 is the actual object, and c is the use of concatenate (combine) 
 function. This could result in conflict errors or have mirroring issues. 
 Overall, it is unwise to use c for the literal object. 

"What kind of data type is “grp”? How would you change it between character
and factor and back again?"
  Grp is a chr type (character). To change this to factor you could write:
```{r}
DF$grp <- as.factor(DF$grp)
```

 To change it back:
```{r}
DF$grp <- as.character(DF$grp)
```


1. Vectors

```{r, error = TRUE}
library(hjl246)
set.seed(1)
v1 <- sample(0:100, 20, replace = TRUE)
v1

v2 <- sample(1:10, 20, replace = TRUE)
v2

v3 <- runif(20, 0, 1)
v3

v4 <- sample(LETTERS[v2], 20)
v4



```

2. Matrices

```{r, error = TRUE}
library(hjl246)
m1 <- matrix(c(v1, v2), ncol = 2)
m1

m2 <- matrix(c(v2, v3), ncol = 2)
m2

m3 <- matrix(c(v3, v4), ncol = 2)
m3
```

3. Data Frame

```{r, error = TRUE}
library(hjl246)
DF <- data.frame(v1, v2, v3, v4)
DF

colnames(DF) <- c("a", "b", "c", "grp")
DF
```

4. 'my_calc' Examples

```{r, error = TRUE}
library(hjl246)
s1 <- my_calc(v1, v2, "sum")
s1
d1 <- my_calc(v1, v2, "divide")
d1
s2 <- my_calc(v2, v3, "sum")
s2
d2 <- my_calc(v2, v3, "divide")
d2
s3 <- my_calc(v1, v3, "sum")
s3
d3 <- my_calc(v1, v3, "divide")
d3
s4 <- my_calc(m1, m1, "sum")
s4
d4 <- my_calc(m1, m1, "divide")
d4
s5 <- my_calc(m1, m2, "sum")
s5
d5 <- my_calc(m1, m2, "divide")
d5


df1 <- my_calc(DF[1, 1], DF[1, 2], "multiply")
df1
df2 <- my_calc(DF[1:10, 1:3], DF[1:10, 1:3], "multiply")
df2
df3 <- my_calc(DF[which(DF[, "grp"] == "E"), c("a", "b")], DF[which(DF[, "grp"] == "E"), "c"], "divide")
df3

DF[which(DF[, "grp"] == "E"), c("a", "b")] / DF[which(DF[, "grp"] == "E"), "c"]
```

5. *apply Examples

```{r, error = TRUE}
library(hjl246)
l = list(v1, m1, m2)
l

l1 <- lapply(l, sum)
l1


l2 <- lapply(1:length(l), function(x) {
    o <- my_calc(l[[x]], v2, "multiply")
    return(o)
})
l2


l3 <- sapply(1:length(l), function(x) {
    o <- my_calc(l[[x]], v2, "multiply")
    return(o)
})
l3
```


