---
title: "unit2-module2"
author: "Harrison Leduc"
date: "2024-03-30"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{unit2-module2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hjl246)
```

DOUBLE CHECK WHICH OF THESE NEED TO BE LOADED IN ##################
##############eERROR##########################################
Data
```{r}
library(geospaar)
chirps <- rast(system.file("extdata/chirps.tif", package = "geospaar"))
districts <- system.file("extdata/districts.geojson", package = "geospaar") %>%   st_read %>% mutate(ID = 1:nrow(.))
chirpsz <- mask(x = chirps, mask = districts)

farmers <- system.file("extdata/farmer_spatial.csv", package = "geospaar") %>%     read_csv(show_col_types = FALSE) 
roads <- system.file("extdata/roads.geojson", package = "geospaar") %>% st_read


```

Task 1
```{r}

districts_ss <- districts[c(22, 26, 53, 54), ]
districts_ss
  
r <- rast(x = ext(districts_ss), crs = crs(districts), resolution = 0.1)
r

set.seed(1)
rsamp <- r
values(rsamp) <- sample(10:50, size = ncell(rsamp), replace = TRUE)
rsamp

rrandn <- r
values(rrandn) <- rnorm(n = ncell(rrandn), mean = 30, sd = 5)
rrandn

#s <- stack(rsamp, rrandn)
#s <- stack(rsamp, rrandn) %>% 
#  mask(x = ., mask = districts_ss)
#s

s <- c(rsamp, rrandn) %>% 
  mask(x = ., mask = districts_ss)

plot_noaxes(s, axes = FALSE, mar = c(1, 0.5, 1, 4))
#s #????

#test_m <- mask(x = chirps[[1]], mask = districts)
#plot(test_m, axes = FALSE, mar = c(1, 0.5, 1, 4)) 

#plot_noaxes(s, axes = FALSE, mar = c(1, 0.5, 1, 4))  

```

Task 2
```{r}
s2_1d <- disagg(x = s[[1]], fact = 4, method = 'bilinear') 
s2_1d


s2_1gt35 <- s2_1d > 35
s2_1gt35

plot(s2_1gt35)
plot(app(s2_1gt35, sum))

  
#s2_1gt35 <- rast(x = s2_1gt35, resolution = 0.025, crs = crs(s2_1d))

s2_1gt35[s2_1gt35 == 0] <- NA

test3 <- as.points(x = s2_1gt35) %>% 
  st_as_sf


test <- s2_1gt35 > unlist(global(s2_1gt35, mean, na.rm = TRUE))


str(test)
plot(s2_1d, main = paste0("Disagg bilinear"))
plot(test, add = TRUE)


```

Task 3
```{r}
zamr <- rast(x = ext(districts), crs = crs(districts), resolution = 0.5)
zamr
values(zamr) <- 1
zamr


#farmers <- system.file("extdata/farmer_spatial.csv", package = "geospaar") %>%     read_csv(show_col_types = FALSE) 

farmersr <- farmers %>% 
  distinct(uuid, .keep_all = TRUE) %>% 
  dplyr::select(x, y) %>% 
  mutate(count = 1) %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326) %>% 
  rasterize(x = ., y = zamr, field = "count", fun = sum) %>% 
  print()
farmersr

farmersr <- farmersr %>% 
  mask(x = ., mask = districts)


par(mar = c(0, 0, 1, 4))
districts %>% 
  st_union %>% 
  plot(col = "grey", border = "grey", 
       main = expression(paste("N farmers per 0.5", degree, " cell")))
plot(farmersr, add = TRUE, axes = FALSE)
```

Task 4
```{r}
farmersrpts <- as.points(x = farmersr) %>% 
  st_as_sf
farmersrpts

zamr <- rast(x = ext(zamr), crs = crs(zamr), resolution = 0.05)
zamr

dist_to_farmers <- distance(x = zamr, y = farmersrpts)
  
  
dist_to_farmers <- dist_to_farmers %>% 
  mask(x = ., mask = districts)


plot_noaxes(dist_to_farmers / 1000)
plot(farmersrpts, pch = 19, col = "black", add = TRUE)
```

Task 5


```{r}
wctemp <- geodata::worldclim_country(var = "tavg", res = 2.5, 
                                     country = "Zambia", path = tempdir())

zamtmean <- mask(app(wctemp, mean), districts)
zamtmean

plot(zamtmean, axes = FALSE, nr = 1, mar = c(1, 0.5, 1, 4))

```

Task 6
```{r}
trng <- global(zamtmean, range, na.rm = TRUE)
reclmat <- cbind(c(floor(trng[1]), 20, 24), c(20, 24, ceiling(trng[2])), 1:3)

zamtclass <- classify(x = zamtmean, rcl = unlist(reclmat), include.lowest = TRUE)

cols <- c("blue", "yellow2", "red")

plot_noaxes(zamtclass, legend = FALSE, main = "Average Temps", col = cols,
            mar = c(1, 1, 1, 3))
legend(x = "bottomright",
       col = cols, pch = 15, pt.cex = 3,  bty = "n",
       legend = c('Low', 'Medium', 'High'))
```

Ceiling and floor are rounding functions. To be specific...

(Formal definitions from the help page)
The floor function takes a single numeric argument and gives a numeric vector
containing the largest integers not greater than the initial single single
numeric argument.

The ceiling function takes a single numeric argument and gives a numeric
vector that holds the smallest integers that aren't less than the initial 
single numeric argument.

(Simple summary)
Floor gives the largest integer that isn't greater than the input. While 
ceiling gives the smallest integer that isn't less than the input.

Task 7
```{r}
wcprec <- geodata::worldclim_country(var = "prec", res = 2.5, 
                                     country = "Zambia", path = tempdir())
zamprec <- mask(app(wcprec, mean), districts)

zamprec_rs <- resample(x = zamprec, y = zamtclass, method = "near")  # match extent
z <- zonal(x = zamprec_rs, z = zamtclass, fun = "mean")


zamprecz <- subst(x = zamtclass, from = z[, 1], to = z[, 2])
zamprecz

cols <- c("yellow2", "green3", "blue")


plot_noaxes(zamprecz, legend = FALSE, main = "Average Rainfall", col = cols, 
            mar = c(1, 1, 1, 3))
legend(x = "bottomright", col = cols, pch = 15, pt.cex = 3,  bty = "n",
       legend = round(z$mean.1))
```

Aggregate it to the same resolution as chirps, using the default mean aggregation, and mask it using districts. Call that dem5. Use terrain to calculate aspect from dem5 (call it aspect), selecting degrees as the output value. Then find all west-facing aspects (aspects >247.5 and <292.5), and all east facing aspects (>67.5 and <112.5), making new rasters respectively named west and east, e.g. west <- aspect > 247.5 & aspect < 292.5). Stack these together with aspect and make a three-panel plot with plot_noaxes with titles “Aspect”, “West”, and “East”. (Ref: Chunks 37, 42)

```{r}
dem <- geodata::elevation_30s(country = "ZMB", path = tempdir())
dem5 <- aggregate(x = dem, fact = 6)
dem5 # resolution is 0.05

dem5 <- mask(x = dem, mask = districts)

aspect <- terrain(x = dem5, v = 'aspect', unit = 'degrees')
west <- (aspect > 247.5 & aspect < 292.5)
east <- (aspect > 67.5 & aspect < 112.5)

s <- c(aspect, west, east)
titles <- c("Aspect", "West", "East")
plot_noaxes(s, main = titles)


```
