---
title: "unit2-module2"
author: "Harrison Leduc"
date: "2024-03-30"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{unit2-module2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hjl246)
```

Data

```{r}
library(geospaar)
districts <- system.file("extdata/districts.geojson", package = "geospaar") %>%   st_read %>% mutate(ID = 1:nrow(.))

farmers <- system.file("extdata/farmer_spatial.csv", package = "geospaar") %>%     read_csv(show_col_types = FALSE) 
```

Task 1
```{r}

districts_ss <- districts[c(22, 26, 53, 54), ]
districts_ss
  
r <- rast(x = ext(districts_ss), crs = crs(districts), resolution = 0.1)
r

set.seed(1)
rsamp <- r
values(rsamp) <- sample(10:50, size = ncell(rsamp), replace = TRUE)
rsamp

rrandn <- r
values(rrandn) <- rnorm(n = ncell(rrandn), mean = 30, sd = 5)
rrandn


par(mar = rep(0, 4))
s <- c(rsamp, rrandn) %>% 
  mask(x = ., mask = districts_ss)

plot_noaxes(s, axes = FALSE, mar = c(1, 0.5, 1, 4))
```

Task 2
```{r}
s2_1d <- disagg(x = s[[1]], fact = 4, method = 'bilinear') 
s2_1d

s2_1gt35 <- s2_1d > 35
s2_1gt35

s2_1gt35[s2_1gt35 == 0] <- NA

s2poly <- as.polygons(x = s2_1gt35) %>% 
  st_as_sf

plot(s2_1d, main = paste0("Disagg bilinear"))
plot(s2poly %>% st_geometry(), add = TRUE)


```

Task 3
```{r}
zamr <- rast(x = ext(districts), crs = crs(districts), resolution = 0.5)
zamr
values(zamr) <- 1
zamr


farmersr <- farmers %>% 
  distinct(uuid, .keep_all = TRUE) %>% 
  dplyr::select(x, y) %>% 
  mutate(count = 1) %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326) %>% 
  rasterize(x = ., y = zamr, field = "count", fun = sum) %>% 
  print()
farmersr

farmersr <- farmersr %>% 
  mask(x = ., mask = districts)


par(mar = c(0, 0, 1, 4))
districts %>% 
  st_union %>% 
  plot(col = "grey", border = "grey", 
       main = expression(paste("N farmers per 0.5", degree, " cell")))
plot(farmersr, add = TRUE, axes = FALSE)
```

Task 4
```{r}
farmersrpts <- as.points(x = farmersr) %>% 
  st_as_sf
farmersrpts

zamr <- rast(x = ext(zamr), crs = crs(zamr), resolution = 0.05)
zamr

dist_to_farmers <- distance(x = zamr, y = farmersrpts)
  
  
dist_to_farmers <- dist_to_farmers %>% 
  mask(x = ., mask = districts)


plot_noaxes(dist_to_farmers / 1000)
plot(farmersrpts, pch = 19, col = "black", add = TRUE)
```

Task 5


```{r}
wctemp <- geodata::worldclim_country(var = "tavg", res = 2.5, 
                                     country = "Zambia", path = tempdir())

zamtmean <- mask(app(wctemp, mean), districts)
zamtmean

plot(zamtmean, axes = FALSE, nr = 1, mar = c(1, 0.5, 1, 4))
```

Task 6
```{r}
trng <- global(zamtmean, range, na.rm = TRUE)
reclmat <- cbind(c(floor(trng[1]), 20, 24), c(20, 24, ceiling(trng[2])), 1:3)

zamtclass <- classify(x = zamtmean, rcl = unlist(reclmat), include.lowest = TRUE)

cols <- c("blue", "yellow2", "red")

plot_noaxes(zamtclass, legend = FALSE, main = "Average Temps", col = cols,
            mar = c(1, 1, 1, 3))
legend(x = "bottomright",
       col = cols, pch = 15, pt.cex = 3,  bty = "n",
       legend = c('Low', 'Medium', 'High'))
```

Ceiling and floor are rounding functions. To be specific...

(Formal definitions from the help page)
The floor function takes a single numeric argument and gives a numeric vector
containing the largest integers not greater than the initial single single
numeric argument.

The ceiling function takes a single numeric argument and gives a numeric
vector that holds the smallest integers that aren't less than the initial 
single numeric argument.

(Simple summary)
Floor gives the largest integer that isn't greater than the input. While 
ceiling gives the smallest integer that isn't less than the input.

Task 7
```{r}
wcprec <- geodata::worldclim_country(var = "prec", res = 2.5, 
                                     country = "Zambia", path = tempdir())
zamprec <- mask(app(wcprec, sum), districts)

zamprec_rs <- resample(x = zamprec, y = zamtclass, method = "near")  # match extent
z <- zonal(x = zamprec_rs, z = zamtclass, fun = "mean")


zamprecz <- subst(x = zamtclass, from = z[, 1], to = z[, 2])
zamprecz

cols <- c("yellow2", "green3", "blue")


plot_noaxes(zamprecz, legend = FALSE, main = "Average Rainfall", col = cols, 
            mar = c(1, 1, 1, 3))
legend(x = "top", col = rev(cols), pch = 15, pt.cex = 3,  bty = "n",
       legend = round(z$sum))
```

Task 8

```{r}
dem <- geodata::elevation_30s(country = "ZMB", path = tempdir())
dem5 <- aggregate(x = dem, fact = 6)
dem5 # resolution is 0.05

dem5 <- mask(x = dem5, mask = districts)

aspect <- terrain(x = dem5, v = 'aspect', unit = 'degrees')
west <- (aspect > 247.5 & aspect < 292.5)
east <- (aspect > 67.5 & aspect < 112.5)

s <- c(aspect, west, east)
titles <- c("Aspect", "West", "East")
plot_noaxes(s, main = titles)
```

Task 9
```{r}
set.seed(1)
west[west == 0] <- NA
east[east == 0] <- NA

randwest <- spatSample(west, size = 100, xy = TRUE, na.rm = TRUE) %>% 
  st_as_sf(coords = c("x", "y"), crs = st_crs(districts))
randwest
  
randeast <- spatSample(east, size = 100, xy = TRUE, na.rm = TRUE) %>% 
  st_as_sf(coords = c("x", "y"), crs = st_crs(districts))
randeast

temp_stats <- bind_rows(
  tibble(temp = terra::extract(zamtmean, randwest)$mean, dat = "West"), 
  tibble(temp = terra::extract(zamtmean, randeast)$mean, dat = "East")
)
temp_stats


bp_theme <- theme(legend.title = element_blank(), 
                  axis.text.x = element_blank(),
                  axis.ticks.x = element_blank(), 
                  panel.grid.major.x = element_blank(), 
                  panel.grid.minor.x = element_blank(), 
                  panel.background = element_rect(fill = "grey95"))
ggplot(temp_stats) +
  geom_boxplot(mapping = aes(y = temp, fill = dat), position = "dodge2") +
  scale_fill_manual(values = c("steelblue", "cadetblue")) + 
  ggtitle("Temp distributions") + xlab(NULL) + ylab("degrees Celsius") + 
  bp_theme
```
